
\item Jobs are sent to a server at a rate of 2 jobs per minute. We will model job arrivals using a homogenous Poisson process. 
\begin{enumerate}
\item What is the probability of receiving more than 3 jobs in a period of one minute?
\item What is the probability of receiving more than 30 jobs in a period of 10 minutes?
\item What is the expected value and the variance of the time between two consecutive arrivals?
\item Compute the probability that the next job does not arrive during the next 30 seconds?
\item Compute the probability that the time till the fourth job arrives exceeds 40 seconds?
\end{enumerate}

\ansfont{
\begin{enumerate}
\item Let $X(t)$ be the number of calls in $t$ minutes, then $X(t) \sim Po(\lambda t)$ where $\lambda=2$ jobs per minute. 
\[ P[X(1)>3]=1-P[X(1) \le 3]=1-0.8571=0.1429\]

\item \[ P[X(10)>30]=1-P[X(10) \le 30]=1-0.9865=0.0135 \]

\item Let $I_j=$ be the inter-arrival time for the $j$th job. Then $I_j \stackrel{iid}{\sim} Exp(2)$ and
\[ E(I_j)=\frac{1}{\lambda}=0.5 \mbox{ min} \qquad Var(I_j)=\frac{1}{\lambda^2}=0.25 \mbox{ min}^2\]

\item Since 30 seconds equals 0.5 minutes, we have 
\[P[I_1>0.5 \mbox{ min}]=e^{-\lambda(0.5)}=e^{-1}=0.3679\]

\item Let $O_4=(I_1+I_2+I_3+I_4)$ be the time until the fourth job, then $O_4 \sim\ Gamma(4,2)$. Now 40 seconds is 2/3 of a minute. We could get this answer directly by using the gamma cumulative distribution function to find $P(O_4>2/3) = 1-P(O_4\le 2/3) = 0.9535$. 

Alternatively, we could use the Poisson-Gamma formula by recognizing the only way the 4th job can take longer than 40 seconds is if, only 0, 1, 2, or 3 jobs arrive in the first 40 seconds (2/3 of a minute). So remembering that $X(2/3) \sim Po(2\times 2/3)$, we can calculate 
\begin{align*}
P(O_4>40/60)&= P(X(2/3) \le 3) \\
&= P(X(2/3) = 0) +  P(X(2/3) = 1) + P(X(2/3) = 2) + P(X(2/3) = 3)\\
&= 0.9539
\end{align*}
\end{enumerate}
}
