\item A small pizza shop has a single telephone number that can ring two 
different phones. Each of these phones has the capability to put a single call 
on hold and then whenever a call is completed the next waiting call (from either 
phone) will automatically ring through. The average call lasts 5 minutes and, 
during peak times, calls come in at a rate of one call every two minutes.
\begin{enumerate}
\item What is the state space for this queuing system?
\item What is the steady-state distribution for this queuing system? Describe
any assumptions you make.
\item The owner of the pizza shop is contemplating adding an additional phone 
line. Do you think the owner should? Why or why not?
\end{enumerate}

\ansfont{
\begin{enumerate}
\item The state space is $\{0,1,2,3,4\}$. 
\item Let's assume this queuing system has independent exponentially distributed 
arrival and service times and therefore can be treated as an $M/M/2/4$ system 
with arrival rate $\lambda_A = 1/2$ calls per minute and service rate $\lambda_S
 = 1/5$ calls per minute. Thus $r=\lambda_A/\lambda_S = (1/2)/(1/5) = 2.5$. First
find 
\[ \begin{array}{rl}
\pi_0 &
= \left( \sum_{x=0}^1 \frac{r^x}{x!} + 
\sum_{x=2}^4 \frac{r^2}{2!}(r/2)^{x-2} \right)^{-1} \\
&
\left(
= \frac{2.5^0}{0!}
+ \frac{2.5^1}{1!} 
+ \frac{2.5^2}{2!}(2.5/2)^{2-2} 
+ \frac{2.5^2}{2!}(2.5/2)^{3-2} 
+ \frac{2.5^2}{2!}(2.5/2)^{4-2} 
\right)^{-1} \\
&= 15.414^{-1} \\
&= 0.065
\end{array} \]
Then 
\[ \begin{array}{rl}
\pi_1 &= \frac{r^1}{1!}\pi_0 = 2.5\pi_0 = 0.162 \\
\pi_2 &= \frac{r^2}{2!}\pi_0 = 2.5^2\pi_0/2 = 0.203 \\
\pi_3 &= \frac{r^k}{k!}(r/k)^{(3-k)}\pi_0 
       = \frac{2.5^2}{2!}(2.5/2)^{(3-2)}\pi_0 = 0.253 \\
\pi_4 &= \frac{r^k}{k!}(r/k)^{(4-k)}\pi_0 
       = \frac{2.5^2}{2!}(2.5/2)^{(4-2)}\pi_0 = 0.317 \\
\end{array} \]
\item Yes since the system is at capacity 31.7\% of the time, the shop is 
missing 31.7\% of the calls or $0.317\lambda_A=0.8$ calls per minute. Adding 
another phone line would reduce the number of missed calls dramatically.
\end{enumerate}
}