\item There is only one restroom in a restaurant, where customers go to this
    restroom according to a
    Possion process with average rate of 16 people every hour. The time that a
    customer uses the restroom follows an exponential distribution with expectation
    3 minutes. If the restroom is being used, newly arrived customers stay in the queue.
    Find steady-state, i.e. large $t$, behavior of the following quantities.
    \begin{enumerate}
      \item The expected number of customers in the restroom and waiting at any time.
      \item The proportion of time when no one is using the restroom.
      \item The probability that there are less than 3 customers either waiting
        or using the restroom.
      \item The average time that an arrival stands in the queue.
      \item The probability that a customer is in the queue for longer than 15 
      min. This can be calculated using the formula $P(W > w) = re^{-w/E[R]}$ 
      where $W$ is the wait time and $R$ is the response time.
    \end{enumerate}
    \ansfont{
      This is an $M/M/1$ queue with utilization $r=\lambda_A/\lambda_S=(16/60)/(1/3)=4/5$.
      Then the steady state probabilities are
      \begin{eqnarray*}
        p_0 &=& 1-4/5=.2\\
        p_1 &=&(.8)(.2)=.16\\
        p_2 &=&(.8)^2(.2)=.128\\
        \vdots &=& \vdots
      \end{eqnarray*}
      \begin{enumerate}
        \item $E[X]=\frac{r}{1-r}=\frac{4/5}{1/5}=4$
        \item $P(X(t)=0)=p_0=0.2$ implies no one using the restroom 20\% of the time.
        \item $P(X(t)<3)=P(X(t)\le 2)=p_0+p_1+p_2=.2+.16+.128=.488$
        \item Expected response time in the system is $E[R]=E[X]/\lambda_A= 4/(4/15)=15$ min and expected service time is $E[S]=1/\mu=1/(1/3)=3$. Thus $E[W]=E[R]-E[S]=5-3=12$ min.
        \item $P(W>15)=(4/5)e^{-15/15}=0.294$.
      \end{enumerate}
    }